---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


This project makes use of [RStudio's projects][Rproj] ("*.Rproj").
The folder, in which the .Rproj-file is located in,
is automatically set as the **working directory** by RStudio
when opening the .Rproj-file.
Thus there is no need to explicitly set the working directory.

Clear the workspace.
```{r}
rm(list = ls())
```

Load the packages used in this project (and install them if they are not yet installed).
The custom `load.packages` function takes a character vector of R-packages names.

```{r}
# Define the `load.packages` function
load.packages <- function(packages) {
    for (package in packages) {
        if (!package %in% installed.packages())
            install.packages(package)
        library(package, character.only = TRUE)
    }
}

# Load the packages
load.packages(c("dplyr",
                "magrittr",
                "readr"))
```


## Loading and preprocessing the data



## What is mean total number of steps taken per day?



## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?




[Rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

