---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


This project makes use of [RStudio's projects][Rproj] ("*.Rproj").
The folder, in which the .Rproj-file is located in,
is automatically set as the **working directory** by RStudio
when opening the .Rproj-file.
Thus there is no need to explicitly set the working directory.

Clear the workspace.
```{r}
rm(list = ls())
```

Load the packages used in this project (and install them if they are not yet installed).
The custom `load.packages` function takes a character vector of R-packages names.

```{r}
# Define the `load.packages` function
load.packages <- function(packages) {
    for (package in packages) {
        if (!package %in% installed.packages())
            install.packages(package)
        library(package, character.only = TRUE)
    }
}

# Load the packages
load.packages(c("dplyr",
                "magrittr",
                "readr"))
```


## Loading and preprocessing the data

Unzip the ZIP-file.

```{r}

# unzip(zipfile = "activity.zip")

```


Read the CSV-file using `read_delim` from the `readr`package.

```{r}
dat <- read_delim(file = "activity.zip", 
                  delim = ",",
                  col_types = list(col_integer(),
                                   col_datetime(),
                                   col_integer()))
```

There are **`r format(nrow(dat), big.mark = ",")` observations** 
in this data set, which is **the 
`r ifelse(nrow(dat) == 17568, "same", "not the same")`**
as the 17,568 observations mentioned in the instructions.

The data types of the three variables are:

```{r}
glimpse(dat)
```

These are the six and last six observations, respectively.

```{r}
head(dat)
tail(dat)
```

Here are summary statistics for the three variables.

```{r}
lapply(dat, summary)
```





## What is mean total number of steps taken per day?



## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?




[Rproj]: https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects

